---
title: "DataAnalysis"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(afex)
library(emmeans)

load("data/USSurvey.Rda")
load("data/SpanishSurvey.Rda")
load("data/fMRIdata.Rda")

```

```{r Demographic variables: education and income, echo = FALSE}

#---------------Spanish survey-----------------
# income (net household monthly income): 1 = no income, 2 = <300 euros/mo,  3 = 300-600, 4 = 600-900, 5 = 900-1200, 6 = 1200-1800, 7 =  1800-2400, 8 = 2400-3000, 9 = 3000-4500, 10 = 4500-6000, 11 = <6000

#education: 1 = Ningún nivel, 2 = Primaria, 3 = Secundaria, 4 = Bachillerat, 5 = FP, 6 = Uni no finalizada, 7 = Diplomatura, 8 = Licenciatura, 9 = Master, 10 = Grado Medio, 11 = Doctorado, 12/13 = NS/NC

#---------------US survey----------------------
#education: 1 = no school, 2 = 8th grade or less, 3 = some high school, 4 = High school or equivalent, 5 = Some college, 6 = 4 year collage degree, 7 = Graduate or professional training

#---------------fMRI study----------------------
#education: 1 = None, 2 = Primaria, 3 = Secundaria, 4 = Bachillerato, 5 = FP, 6 = Grado universitario, 7 = Master/posgrado, 8 = Doctorado

##income (net household monthly income): 1 = < 900, 2 = 900 - 1200, 3 = 1200- 2000, 4 = 2000-3000, 5 = 3000-4500, 6 = 4500-6000, 7 = < 6000, 8 = NC

```


## SACRED VALUES

# How did sacred values influence misinformation sharig?

Misinformation that included sacred values was *more likely to be shared* across the three samples.


```{r effect of sacred values}

#-------------------------------------------------
#               Spanish survey
#-------------------------------------------------
SacredFit <- ESdf %>% 
  group_by(identity, value) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value + (1|identity), .) 

summary(SacredFit)
PairwiseComparison <- emmeans(SacredFit, pairwise ~ value)
PairwiseComparison

#-------------------------------------------------
#               US survey
#-------------------------------------------------
SacredFit <- USdf %>% 
  group_by(identity, value) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value + (1|identity), .) 

summary(SacredFit)
PairwiseComparison <- emmeans(SacredFit, pairwise ~ value)
PairwiseComparison

#-------------------------------------------------
#               fMRI study
#-------------------------------------------------
SacredFit <- MRdf %>% 
  filter(!(identity %in% "1823")) %>% # filter fMRI participant with excessive head movement
  group_by(identity, value) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value + (1|identity), .)

summary(SacredFit)
PairwiseComparison <- emmeans(SacredFit, pairwise ~ value)
PairwiseComparison

```


# Number of sacred values moderated the effect of sacred values.

The more sacred values people held (out of the three sacred values we proposed about immigration, nation and traditional family values), *the more they shared* misinformation about sacred values (vs. non-sacred values).


```{r moderation of number of sacred values on effect of sacred values}

#-------------------------------------------------
#               Spanish survey
#-------------------------------------------------
NumSacredFit <- ESdf %>% 
  group_by(identity, value) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value * NumOfSacredValues + (1|identity), .) 

summary(NumSacredFit)

#-------------------------------------------------
#               US survey
#-------------------------------------------------
NumSacredFit <- USdf %>% 
  group_by(identity, value, NumOfSacredValues) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value * NumOfSacredValues + (1|identity), .) 

summary(NumSacredFit)

#-------------------------------------------------
#               fMRI study
#-------------------------------------------------
NumSacredFit <- MRdf %>% 
  filter(!(identity %in% "1823")) %>% # filter fMRI participant with excessive head movement
  group_by(identity, value, NumOfSacredValues) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value * NumOfSacredValues + (1|identity), .) 

summary(NumSacredFit)

```


# Analytical thinking moderated the effect of sacred values.

Analytical thinking was associated with reduced misinformation sharing for posts that tackled non-sacred values, but *did not affect* posts that tackled sacred values


```{r moderation of congitive reflection on the effect of sacred values}

#-------------------------------------------------
#               Spanish survey
#-------------------------------------------------
# Number of correct responses in the Cognitive reflection test (CRT) moderated the effect of sacred values, so that more correct responses predicted reduced sharing of misinformation related to non-sacred values but was unrelated to sharing of misinformation related to sacred values.
CrtFit <- ESdf %>% 
  group_by(identity, value, crt_correct) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value * crt_correct + (1|identity), .) 

summary(CrtFit)
PairwiseComparison <- emmeans(CrtFit, pairwise ~ crt_correct | value, at = list("crt_correct" = c(0, 3))) #compare no correct responses (0/3 responses) to all correct responses (3/3 responses)
PairwiseComparison

#-------------------------------------------------
#               US survey
#-------------------------------------------------
# the CRT was not administered in the US survey

```


## IDENTITY FUSION

# Identity fusion increased misinformation sharing.

Identity fusion with any of the conservative parties & Political affiliation with extreme options (far-right in Spain) was associated with *increased sharing* of posts with sacred values (vs. non-sacred values).


```{r effect of identity fusion & political affiliation}

#-------------------------------------------------
#               Spanish survey
#-------------------------------------------------
## political affiliation with the far-right (vs. center-right) increased sharing

AffilFit <- ESdf %>% 
  group_by(identity, votersGroup) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lm(mean_sharing ~ votersGroup, .) 

summary(AffilFit)
PairwiseComparison <- emmeans(AffilFit, pairwise ~ votersGroup)
PairwiseComparison

## identity fusion with the far-right increased sharing

FusedFit <- ESdf %>% 
  filter(votersGroup %in% "far-right") %>% 
  group_by(identity, Fused) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lm(mean_sharing ~ Fused, .) 

summary(FusedFit)
PairwiseComparison <- emmeans(FusedFit, pairwise ~ Fused)
PairwiseComparison

## identity fusion with the center-right increased sharing

FusedFit <- ESdf %>% 
  filter(votersGroup %in% "center-right") %>% 
  group_by(identity, Fused) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lm(mean_sharing ~ Fused, .) 

summary(FusedFit)
PairwiseComparison <- emmeans(FusedFit, pairwise ~ Fused)
PairwiseComparison

#-------------------------------------------------
#               US survey
#-------------------------------------------------
## identity fusion with Trump increased sharing

FusedFit <- USdf %>% 
  group_by(identity, FusedTrump) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lm(mean_sharing ~ FusedTrump, .) 

summary(FusedFit)
PairwiseComparison <- emmeans(FusedFit, pairwise ~ FusedTrump)
PairwiseComparison

## identity fusion with the Republican party increased sharing (excluding conservatives fused with Trump)
FusedFit <- USdf %>% 
  filter(FusedTrump %in% F) %>% 
  group_by(identity, FusedRepublican) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lm(mean_sharing ~ FusedRepublican, .) 

summary(FusedFit)
PairwiseComparison <- emmeans(FusedFit, pairwise ~ FusedRepublican)
PairwiseComparison

#-------------------------------------------------
#               fMRI study
#-------------------------------------------------
# Only 5 participants were fused with the far-right in the fMRI study, making it difficult to run any trustworthy analyses

```


# Identity fusion with extreme political options (Trump supporters) moderated effect of sacred values.

Identity fusion with Trump in the US was associated with *increased sharing* of posts including sacred values (vs. non-sacred values).


```{r moderation of identity fusion on effect of sacred values}

#-------------------------------------------------
#               Spanish survey
#-------------------------------------------------
## identity fusion with the far-right did not affect sharing of posts with sacred values (vs. non-sacred values)

ValueFusedFit <- ESdf %>% 
  filter(votersGroup %in% "far-right") %>% 
  group_by(identity, value, Fused) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value * Fused + (1|identity), .) 

summary(ValueFusedFit)
PairwiseComparison <- emmeans(ValueFusedFit, pairwise ~ value | Fused)
PairwiseComparison


## no interaction with value was found with fusion with the center-right
ValueFusedFit <- ESdf %>% 
  filter(votersGroup %in% "center-right") %>% 
  group_by(identity, value, Fused) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value * Fused + (1|identity), .)  

summary(ValueFusedFit)
PairwiseComparison <- emmeans(ValueFusedFit, pairwise ~ value | Fused)
PairwiseComparison

#-------------------------------------------------
#               US survey
#-------------------------------------------------
## identity fusion with Trump increased sharing of posts with sacred values (vs. non-sacred values)

ValueFusedFit <- USdf %>% 
  group_by(identity, value, FusedTrump) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value * FusedTrump + (1|identity), .)  

summary(ValueFusedFit)
PairwiseComparison <- emmeans(ValueFusedFit, pairwise ~ value | FusedTrump)
PairwiseComparison

## however, no interaction with value was found with fusion with the Republican party
ValueFusedFit <- USdf %>% 
  filter(FusedTrump %in% F) %>% 
  group_by(identity, value, FusedRepublican) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ value * FusedRepublican + (1|identity), .)  

summary(ValueFusedFit)
PairwiseComparison <- emmeans(ValueFusedFit, pairwise ~ value | FusedRepublican)
PairwiseComparison

```


## FACT-CHECKING

# Fact-checking only had a small effect in non-fused samples.

Fact-checking did not have any effect on misinformation sharing in the Spanish sample. The group that was exposed to the Twitter fact-check showed a slight reduction in sharing when analysed separately. Fact-checks had some effect in participants who were not fused with Trump in the US survey, but had no effect for those fused with Trump.


```{r effect of fact-checking}

#-------------------------------------------------
#               Spanish survey
#-------------------------------------------------
## factchecking had no effect on sharing when compared to control trials (p = 0.08). However, there is an increase in sharing in fact-check trials compared to baseline. In the US survey we investigate whether this increase from baseline is a temporal effect or a backfire effect from fact-checking by including a control group with no fact-checks.
FactCheckFit <- ESdf %>% 
  group_by(identity, factcheck) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ factcheck + (1|identity), .)

summary(FactCheckFit)

## type of fact-check (Twitter, Accuracy, Literacy) did not moderate the effect of fact-checked trials compared to control trials
FactCheckFit <- ESdf %>% 
  group_by(identity, factcheck, FactCheckGroup) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ factcheck * FactCheckGroup + (1|identity), .)

summary(FactCheckFit)
PairwiseComparison <- emmeans(FactCheckFit, pairwise ~ factcheck |  FactCheckGroup)
PairwiseComparison

# there is a very small effect of the Twitter fact-check when analysed separately
FactCheckFit <- ESdf %>% 
  filter(FactCheckGroup %in% "Twitter") %>% 
  group_by(identity, factcheck) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ factcheck + (1|identity), .)

summary(FactCheckFit)

#-------------------------------------------------
#               US survey
#-------------------------------------------------
## Participants in the control group (no fact-checks) showed an increase in sharing in the experimental block compared to the baseline block, while participants in the experimental group (exposed to fact-checks in the experimental block) showed no increase in sharing in the experimental block compared to baseline.

# In this analysis, we excluded control trials in experimental participants, in order to directly compare control trials (vs. baseline) in control participants with fact-checked trials (vs. baseline) in experimental participants.
FactCheckFit <- USdf %>% 
  filter(!(Group %in% "fact-checked" & factcheck %in% "control")) %>% 
  group_by(identity, Group, block) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ Group * block + (1|identity), .)

summary(FactCheckFit)
PairwiseComparison <- emmeans(FactCheckFit, pairwise ~ block | Group)
confint(PairwiseComparison)


```

# Fusion with Trump moderated the effect of fact-checks in the US sample.

```{r moderation of identity fusion on effect of fact-checks}

#-------------------------------------------------
#               US survey
#-------------------------------------------------
## The fact-check was particularly ineffective for participants fused with Trump, who showed no change in sharing in fact-checked compared to the control trials, while some change was observed in non-fused participants.
FactCheckFit <- USdf %>% 
  filter(Group %in% "fact-checked") %>% 
  group_by(identity, factcheck, FusedTrump) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ factcheck * FusedTrump + (1|identity), .)

summary(FactCheckFit)
PairwiseComparison <- emmeans(FactCheckFit, pairwise ~ factcheck | FusedTrump)
PairwiseComparison

```

# Sacred values did not moderate the effect of fact-checks.

```{r moderation of sacred values on effect of fact-checks}

#-------------------------------------------------
#               Spanish survey
#-------------------------------------------------
## sacred values do not affect participant's response to fact-checks
FactCheckFit <- ESdf %>% 
  group_by(identity, factcheck, value) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ factcheck * value + (1|identity), .)

summary(FactCheckFit)

#-------------------------------------------------
#               US survey
#-------------------------------------------------
## sacred values do not affect participant's response to fact-checks
FactCheckFit <- USdf %>% 
  filter(Group %in% "fact-checked") %>%
  group_by(identity, factcheck, value) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ factcheck * value + (1|identity), .)

summary(FactCheckFit)

```


## OTHER EFFECTS: 

Moral-emotional language had mixed effects across samples. 


```{r effect of moral-emotional language}

#-------------------------------------------------
#               Spanish survey
#-------------------------------------------------
## moral-emotional language slightly increased sharing in the Spanish sample
LanguageFit <- ESdf %>% 
  group_by(identity, language) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ language + (1|identity), .)

summary(LanguageFit)

## we thought language would affect non-sacred values to a greater extent compared to sacred values, because sacred values are highly salient regardless of the wording, but the interaction between language and value was non-significant (p = 0.11)
LanguageFit <- ESdf %>% 
  group_by(identity, language, value) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ language * value + (1|identity), .)

summary(LanguageFit)

#-------------------------------------------------
#               US survey
#-------------------------------------------------
## Moral-emotional language did not have an effect on sharing in the US sample
LanguageFit <- USdf %>% 
  group_by(identity, language) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ language + (1|identity), .)

summary(LanguageFit)

# However, participants fused with Trump were more responsive to language, increasing sharing of items that included moral-emotional language
LanguageFit <- USdf %>% 
  group_by(identity, language, FusedTrump) %>% 
  summarise(mean_sharing = mean(sharing, na.rm = T)) %>% 
  lmer(mean_sharing ~ language * FusedTrump + (1|identity), .)

summary(LanguageFit)


```

